A Method for the spectroscopic inference of fundamental stellar parameters

### 2014-05-14 tex/paper.tex
halfway done addressing first round of SA edits

\newcommand{\prob}{{\rm prob}}
\newcommand{\qN}{\{q_i\}_{i=1}^N}
\newcommand{\qM}{\{q_{im}\}_{i=1,m=0}^{N,M}}
\newcommand{\yN}{\{y_i\}_{i=1}^N}
\newcommand{\vt}{\vec{\theta}}
\newcommand{\vg}{\vt_{\star, {\rm grid}}}
\newcommand{\vpp}{\vt_{\star, {\rm post}}}
\newcommand{\vstar}{\vt_{\star}}
\newcommand{\finst}{f_{\lambda, {\rm inst}}}
\newcommand{\fsynth}{f_{\lambda, {\rm synth}}}
\newcommand{\vN}{\vt_{\rm N}}
\newcommand{\vc}{\vec{c}}
\newcommand{\fM}{ {\bm M}}
\newcommand{\fMi}{M_i}
\newcommand{\fD}{ {\bm D}}
\newcommand{\fDi}{D_i}
\newcommand{\fR}{ {\bm R}}
\newcommand{\dd}{\,{\rm d}}
\newcommand{\trans}{\mathsf{T}}
\newcommand{\Z}{[{\rm Fe}/{\rm H}]}
\newcommand{\A}{[\alpha/{\rm Fe}]}
\newcommand{\matern}{Mat\'{e}rn}

\newcommand{\todo}[1]{ \textcolor{Blue}{\\TODO: #1}}

@@

\section{Method}
Typically, the stellar properties we most desire are effective temperature
 $T_{\rm eff}$, radius or surface gravity $\log g$, and metallicity $\Z$.
Several high-quality libraries of synthetic spectra now exist (PHOENIX, Kurucz,
 more being developed for GAIA) and are parameterized by these fundamental
 stellar parameters, which together, we call $\vg$. 
Synthetic libraries provide high-resolution spectra which span a range of $\vg$
 covering the main sequence of spectral types. The library is typically
 specified on a grid of equal spacing in $T_{\rm eff}$, $\log g$, and $\Z$.

In addition to these fundamental parameters, a star also has several observed
 properties that are a function of its kinematics, geometry, and location in our
 galaxy: projected rotational velocity $v \sin i$, line of sight velocity $v_z$,
 total extinction $A_V$, and solid angle $\Omega$. 
We call these secondary parameters $\vpp$, because we can model their effects
 on the synthetic spectra in a ``post-processing'' step by convolution with an
 appropriate kernel or multiplication with a smooth function. 
Together, we call $\vstar = \{\vg, \vpp\}$.

Using a synthetic stellar library as a backend and applying post-processing
 techniques, we attempt to forward-model the observed stellar spectra in order
 to determine the best-fit $\vstar$. 
Because properly sampled spectra have correlated pixels, we present a framework
 which self-consistently models these correlations.

\begin{table}[!htb]
\begin{tabular}{ll}
\hline
\hline
Symbol & Description\\
\hline
\hline
$i$ & index specifying a pixel\\
$\lambda_i$ & wavelength corresponding to a given pixel $i$\\
$\vg$ & fundamental stellar parameters, $T_{\rm eff}, \log(g), \Z, \A$\\
  & that parameterize a synthetic spectrum from the grid\\
$\vpp$ & stellar parameters $v \sin i$, $v_z$, $A_V$, and $R^2/d^2$ that\\
  & are applied during ``post processing'' of the synthetic spectrum\\
$\vstar$ & $\{\vg,\vpp \}$\\
$\finst(\lambda)$ & data spectrum\\
$\fsynth(\lambda)$ & synthetic spectrum\\
$c_0, c_1, \ldots$ & Chebyshev polynomial coefficients for residual flux calibration\\
$\vN$ & the set of nuisance parameters composed of $\{c_0, c_1, \ldots, c_N\}$\\
$\vt$ & the parameters $\{\vg, \vpp, \vN\}$ that completely describe a model spectrum\\
$\fDi$ & data flux for a given pixel, $D(\lambda_i)$\\
$\fD$ & data vector comprised of all $\fDi$, $i = \{0, \ldots, N\}$\\
$\fMi$ & model flux for a given pixel, $M(\lambda_i | \vt)$\\
$\fM$ & model vector comprised of all $\fMi$, $i = \{0, \ldots, N\}$\\
$\sigma_i$ & Poisson noise for a given pixel $i$\\
$R_i$ & residuals $\fDi - \fMi$\\
$\fR$ & residual vector $\fD - \fM$\\
$C$ & covariance matrix\\
$k_{\rm global}$  & global covariance kernel\\
$k_{\rm region}$  & regional covariance kernel\\
\hline
\end{tabular}
\caption{Nomenclature used in this document}
\label{tab:nomenclature}
\end{table}

\begin{figure*}[!htb]
\begin{center}
  \includegraphics{figs/model_data}
  \caption{\textbf{Top}: Chebyshev polynomial which modifies the model spectrum
   to account for inaccuracies in the flux-calibration. 
  \textbf{Middle}: The data spectrum and model spectrum, after it has been
   interpolated, post-processed, and multiplied by the Chebyshev polynomial.
  \textbf{Bottom}: Residuals from the model fit. Note the large residual at
   $\lambda5189$\AA\ due to a missing opacity source from \ion{Ca}{1}. 
  \protect \todo{Model spectrum will be updated once I properly burn in the
   chain. Do these line styles look OK? I was going for something legible
   that might also print well in B/W}.}
\label{fig:model_data}
\end{center}
\end{figure*}

\subsection{Interpolation from synthetic spectral libraries}
Creating a new spectrum with a specific $\vg$ requires either synthesizing a
 new spectrum using radiative transfer codes or interpolating from nearby grid
 points in the library. 
Because it is computationally expensive to synthesize a spectrum at high
 resolution over a wide wavelength range, we choose to interpolate. 
Spline interpolation for spectra \citep{hus12}.
\todo{Implement band-limited interpolation from the synthetic grid using spline interpolation. Caution: As is, 100K/0.5 logg spacing in $\vg$ may not be Nyquist sampling.}
\todo{Investigate a ``Bayesian Emulator'' for using Gaussian processes for interpolating from a simulation grid.}

\subsection{Post-processing synthetic spectra to match the data}
Libraries of synthetic spectra are generally computed at high resolution ($R
 \geq $100,000), sampled with many pixels per resolution element, and do not
 include any rotational broadening or account for any instrumental effects. 
In order to transform a raw synthetic spectrum interpolated from the grid into
 one that matches the spectrum of a real star, we must post-process the spectrum
 to account for these secondary effects. 
The projected rotational velocity of the star, parameterized by $v \sin i$,
 broadens spectral line profiles and is mathematically described by convolution
 with a kernel $g_{v \sin i}(\lambda)$.  
UV, optical, and infrared spectra are acquired using a spectrograph which
 disperses light onto a CCD, defining a specific resolution and sampling rate
 for the spectrum. 
Because the resolution of the spectrograph and spacing of the pixels are
 different from the synthetic spectral library, we must convolve the raw
 spectrum with the line spread function (LSF) of the instrument $g_{\rm
 LSF}(\lambda)$ and resample to the exact pixels of the CCD. In wavelength
 space, the $v \sin i$ and LSF operations are represented by the convolution of
 the synthetic spectrum with these two kernels
\begin{equation}
  \finst(\lambda) = g_{v \sin i}(\lambda) \otimes g_{\rm LSF}(\lambda) \otimes \fsynth(\lambda)
  \label{eqn:convolution}
\end{equation}
Using the convolution theorem, we can rewrite these operations as
 multiplications in Fourier space 
\begin{equation}
  F_{\lambda, {\rm inst}}(s) = G_{\rm LSF}(s) G_{v \sin i}(s) F_{\lambda, {\rm synth}}(s)
\end{equation}
 where $G \leftrightarrow g$ and $F \leftrightarrow f$ are Fourier transform
 pairs. 
In order to use the fast Fourier transform (FFT) to execute these
 operations using discrete samples of the synthetic spectrum $\fsynth$, we must
 first resample $\fsynth$ so that it is on a uniform grid. 
In the case of a spectrum, the natural uniform grid is one that is equally
 sampled in velocity space, such there is an equal velocity shift $\Delta v$
 between pixels. This results in a wavelength grid that is linearly sampled in
 the logarithm of wavelength. 
Therefore, the Fourier coordinate $s$ is the number of cycles per sampling
 interval, having units of inverse velocity $[{\rm s}/{\rm km}]$. 
Next, we multiply the Fourier transform of the synthetic spectrum by the
 Fourier transforms of the rotational velocity and line spread function kernels.
Lastly, we do an inverse FFT to transform the modified spectrum $F_{\lambda,
 {\rm inst}}$ back to wavelength space, where it is sampled at the wavelength
 locations of the pixels in the detector ($\finst$). 
When resampling the spectrum, we are careful to ensure band-limited
 interpolation using splines in order to prevent introducing non-physical high
 frequency structure into the spectrum.

Finally we apply corrections for interstellar extinction and Doppler shift.
Interstellar extinction attenuates the spectrum by an amount $A_\lambda$, which
is wavelength-dependent. 
We use the (name here) dust-extinction model, which is parameterized by
 extinction in the $V$ photometric band, $A_V$. 
We correct for a radial velocity difference $v_z$ between the star and the
 earth by Doppler shifting the model spectrum. 
\todo{Doppler shifting the wavelengths could be implemented as a phase-shift in
 Fourier space, but it's pretty fast as-is.}

\todo{We might try implementing the convolution in wavelength space, rather
 than the FFT. 
This would allow the LSF to change with wavelength.\\} 
Pixel width effects are not important as long as the LSF of the instrument is
 adequately sampled ($\gtrsim 3$ pixels across the LSF). 
If our spectrum were not properly sampled, we would need to additionally
 convolve with a boxcar the width of the pixels.

\begin{figure*}[!htb]
\begin{center}
  \includegraphics[width=0.9\textwidth]{figs/badlines}
  \caption{The variety of residual behaviour, depending on the quality of model
    fit. 
    From left to right: \textbf{Class 0} covariance results from slight
     model mismatch, and correlates nearby residuals. 
    \textbf{Class I}: A missing absorption line in the model leaves a large,
     highly correlated patch of negative residuals. 
    \textbf{Class II}: An extraneous line in the model leaves a large, highly
     correlated patch of positive residual.  
    \textbf{Class III}: If lines are present in the model but of the wrong
     strength, many correlated residuals of moderate amplitude will result. 
    The difficulty with class III lines is that for any specific line, there might
     exist a $\vstar$ that will fit the line, but there does not exist a
     $\vstar$ that will properly fit \emph{all} the lines.}
\label{fig:badlines}
\end{center}
\end{figure*}

Due to small imperfections in the flux-calibration of the data spectrum, there
 may be broad regions of the spectrum which are offset from the true continuum
 level by a slight amount. 
A traditional approach to address this problem is to normalize both the data
 and the spectrum to the continuum level before comparison. 
While this may work well for hotter stars with well-defined continuua, for
 cooler stars with large molecular features, normalization may introduce
 artificial features into the spectrum when the pseudo-continuum is incorrectly
 placed. 
Instead, we choose to multiply the model spectrum by a low-order Chebyshev
 polynomial that is normalized to unity. 
The coefficients of the polynomial, $c_0, c_1, \ldots$, will be free-parameters
 in our model that we will solve for. 
If the behavior of the spectrograph is characterized by observations of
 spectrophotometric standards, then we can put priors on the coefficients that
 will prevent overfitting of the polynomials. 
For an example of a typical spectrum, and Chebyshev polynomial, see
 Figure~\ref{fig:model_data}.

With more uncertainty, the polynomial multiplication can also be used as a
 substitute for flux-calibration since the sensitivity function of an instrument
 is usually a smooth function with wavelength. 
When spectra are not flux calibrated, there is greater uncertainty about the
 properties of the spectrograph and wider priors must be used on the Chebyshev
 coefficients. 
In this case, the polynomial might destroy broad-scale information present in
 the spectrum by fitting it out, something that would be avoided with more
 accurate priors provided by the flux calibration for the model.

\subsection{Likelihood function}
To evaluate which parameters of the model $\fM$ fit the data set $\fD$ best we
 use a standard multidimensional Gaussian likelihood function which allows for
 covariance between data points. 
If the vector of residuals is 
\begin{equation}
  \fR(\vstar) = \fD - \fM(\vstar)
\end{equation}
 with a length of $N$ data points, then the likelihood function is
\begin{equation}
  p(\fD | \vstar) = \frac{1}{\sqrt{(2 \pi)^N \det(C)}} \exp\left ( -\frac{1}{2} \fR^T C^{-1} \fR \right ) 
\end{equation}
 and the log-likelihood function is
\begin{equation}
    \ln[p(\vstar)] = -\frac{1}{2} \fR^T C^{-1} \fR - \frac{1}{2} \ln \det C  - \frac{N}{2} \ln 2 \pi 
    \label{eqn:lnprob}
\end{equation}
$C$ is the covariance matrix which describes the covariance between pixels in
 the spectrum. 
For independent noise with noise per pixel $\sigma_i$, this matrix is diagonal
\begin{equation}
  C = 
  \begin{bmatrix}
    \sigma_0^2 & 0  & \cdots & 0\\
    0 & \sigma_1^2 & \cdots & 0\\
    \vdots  &   & \ddots  & \vdots \\
    0 & 0 & \cdots & \sigma_N^2\\
  \end{bmatrix}
  \label{eqn:covariance_diagonal}
\end{equation}
with $\sigma_{ij} = 0$ everywhere and equation~\ref{eqn:lnprob} reduces to the
 familiar $\chi^2$ form of a sum over the square of the residuals, weighted by
 the inverse variance of each data point
\begin{equation}
  \ln[p(\vstar)] \propto - \frac{1}{2} \chi^2 = - \frac{1}{2} \sum_i \frac{R_i^2}{\sigma_i^2}
\end{equation}

However, residuals from a spectroscopic fit are often correlated due to
 \todo{DUE TO WHAT, exactly?? I'm confused. 
Wait for response from Tom Loredo}. 
Additionally, systematic errors in the synthetic spectra (such as incorrect
 line strengths) will result in regions of highly correlated residuals. 
Properly accounting for correlations in the residuals requires that we use a
 covariance matrix with off-diagonal terms $\sigma_{ij}$ and a likelihood
 function (Equation~\ref{eqn:lnprob}) which uses a matrix product instead of a
 sum over independent pixels.

We seek to understand and parameterize two types of covariance structure in our
 model, exemplified in Figure~\ref{fig:badlines}: the large scale but generally
 mild global covariance structure that results from [still to be finalized
 cause], and a regional covariance structure that is localized to areas where
 spectral line mismatch results in small patches of highly correlated residuals. 
Ignoring the complexity of spectroscopic residuals will bias the estimates of
 the model parameters $\vt$.

As a simple analogy, consider the fit of a straight line to a dataset. 
If the noise in the dataset is correlated, then adjacent data points might be
 offset from the linear trend in the same direction by a similar amount. 
If the covariant structure of the noise is ignored, a simple $\chi^2$ will
 treat these correlated offsets as part of the linear trend, which will result
 in a biased determination of the slope and intercept of the line, typically
 with uncertainties that are too small.

\subsection{Parameterizing the covariance structure}
We parameterize the covariance structure using covariance kernels, which sets
 the covariance between two pixels $\lambda_i$ and $\lambda_j$. 
This behavior is analogous to the two-point correlation function used in
 cosmology, where the distance between two galaxies is used instead of
 wavelength.

\paragraph{Global covariance structure}
A global covariance structure may result from a slight mismatch in the
 continuum or mild mismatch of spectral lines like that shown as class 0 in
 Figure~\ref{fig:badlines}. 
To account for this structure, which is generally mild and typically exists
 across only a few pixels, we use a \emph{stationary} covariance kernel, which
 satisfies the property that the degree of correlation only depends on the
 distance between the two pixels $r$. 
Functions like this are also called \emph{radial basis functions}. 
For spectra, we map $r$ to the velocity difference between two pixels
\begin{equation}
  r(\lambda_i, \lambda_j) = \Delta v = \frac{c}{2} \left | \frac{\lambda_i - \lambda_j}{ \lambda_i + \lambda_j} \right |
\end{equation}
We choose the \matern\ $3/2$ kernel because\ldots
\todo{Explore difference between using \matern\ kernel and squared exponential
 kernel. 
Why \matern\ and why Hann? 
Because they are the vanilla covariance functions that are tried and tested in
 the machine learning communities.}
\begin{equation}
  k_{\nu = 3/2}(r, a, l) = a \left(1 + \frac{\sqrt{3} r}{l} \right ) \exp \left (- \frac{\sqrt{3} r}{l} \right )
\end{equation}
In order to keep the covariance matrix sparse for computational efficiency, we
 taper for compact support using a Hann window
\begin{equation}
  w(r, r_0) = \frac{1}{2} + \frac{1}{2} \cos \left( \frac{\pi r}{r_0} \right) 
\end{equation}
for a global covariance kernel of 
\begin{equation}
  k_{\rm global} (r, a, l, r_0) = w(r, r_0) k_{\nu = 3/2}(r, a, l) 
  \label{eqn:global}
\end{equation}
We add this covariance kernel to the independent variance already present along
 the matrix diagonal
\begin{equation}
  k(\lambda_i, \lambda_j) = k_{\rm global}(r, a, l, r_0) + \delta_{ij} \sigma_{ij}^2
\end{equation}
where $\delta_{ij}$ is the Kroenecker delta function.

\begin{figure*}[!htb]
\begin{center}
\includegraphics[width=0.4\textwidth]{figs/matern_matrix}
\includegraphics[width=0.4\textwidth]{figs/matern_draw}
\caption{\textbf{Left} a covariance matrix generated with the \matern\ kernel
 and typical parameters for our dataset.  
 \textbf{Right} general spectroscopic residuals near a stellar continuum
  overlaid with a sample draw using the covariance matrix, showing that the two
  look similar in structure and amplitude. 
  \protect \todo{Prettify these figures. Consistent labelling. Change
  color stretch so Matrix zeros are white.}}
\label{fig:matern}
\end{center}
\end{figure*}

>>>>

\newcommand{\prob}{{\rm prob}}
\newcommand{\qN}{\{q_i\}_{i=1}^N}
\newcommand{\qM}{\{q_{im}\}_{i=1,m=0}^{N,M}}
\newcommand{\yN}{\{y_i\}_{i=1}^N}
\newcommand{\vt}{\vec{\theta}}
\newcommand{\vg}{\vt_{\star, {\rm grid}}}
\newcommand{\vpp}{\vt_{\star, {\rm post}}}
\newcommand{\vstar}{\vt_{\star}}
\newcommand{\finst}{f_{\lambda, {\rm inst}}}
\newcommand{\fsynth}{f_{\lambda, {\rm synth}}}
\newcommand{\vN}{\vt_{\rm N}}
\newcommand{\vc}{\vec{c}}
\newcommand{\fM}{ \vec{{\bm M}}}
\newcommand{\fMi}{M_i}
\newcommand{\fD}{ \vec{{\bm D}}}
\newcommand{\fDi}{D_i}
\newcommand{\fR}{ {\bm R}}
\newcommand{\dd}{\,{\rm d}}
\newcommand{\trans}{\mathsf{T}}
\newcommand{\Z}{[{\rm Fe}/{\rm H}]}
\newcommand{\A}{[\alpha/{\rm Fe}]}
\newcommand{\matern}{Mat\'{e}rn}

\newcommand{\todo}[1]{ \textcolor{Blue}{\\TODO: #1}}
\newcommand{\comm}[1]{ \textcolor{Red}{SA: #1}}

@@

\section{Method}

A single spectrum provides a wealth of information that can be used to infer the fundamental properties of a star: effective temperature $T_{\rm eff}$, surface gravity $\log g$, and metallicity $\Z$; together $\vg$. We wish to extract the maximal amount information from a given spectrum while also recognizing the limitations imposed by our stellar model and exposing any degeneracies between the parameters.  For example, if we model a large range of the spectrum at once (e.g., the full optical range at high resolution), then we must explicitly account for any covariances and biases introduced by improper calibration or inaccuracies in the model spectrum. We present a generative model which uses a synthetic stellar library as a backend and apply post-processing techniques to mimic the observed spectrum as a function of several stellar parameters and calibration parameters in order to determine the best fit. By forward modelling the data spectrum in a Bayesian framework, our modular approach can easily incorporate additional calibration artifacts and stellar complexity (e.g., veiling or Zeeman splitting) while still being true to the limitations of the dataset.  This data driven approach will enable us to learn how the underlying synthetic models may be improved and refined in an iterative manner.

The roadmap for our method is as follows.
 First, for a given set of fundamental stellar parameters $\vg$, we generate a synthetic
 spectrum by interpolating from a grid of synthetic spectra (\S\ref{subsec:synthetic}). 
Then, we post-process the interpolated spectrum for realism by using additional
 parameters to describe Doppler shift, projected rotational velocity,
 interstellar extinction, and residual flux calibration errors
 (\S\ref{subsec:postprocess}). 
We perform a pixel-by-pixel comparison of the generated spectrum with the
 data spectrum using a prescribed likelihood function
 (\S\ref{subsec:likelihood}) and a parametric treatment of the covariance
 between pixel residuals (\S\ref{subsec:covariance}). 
We use a Markov chain Monte Carlo (MCMC) Gibbs sampler to numerically explore
the posterior probability density and determine constraints on the stellar
and nuisance parameters (\S\ref{subsec:MCMC}). 
The benefit of our method is that it provides robust estimates of the stellar and nuisance
parameters with well-quantified uncertainties and parameter degeneracies. 
Our parameterization of the covariance between pixel residuals also yields a means to identify
 and quantify the key discrepancies between the data and model, and provides a path forward to 
 improve the stellar models in an iterative fashion (\S\ref{subsec:learning}).
 \todo{I really want to overload the word ``covariance'' and use it to describe both the covariance between $T_{\rm eff}$ and $\log g$ as well as the covariance structure of the pixel residuals themselves. However I think this might be confusing so I am using ``degeneracies'' for the former and reserving ``covariance'' for the later.}

\subsection{Synthetic Model Spectra}
\label{subsec:synthetic}
Many methods exist to generate a synthetic model of a stellar spectrum. Generally, when
a high quality spectrum is needed for a specific set of stellar parameters, one
runs a stellar atmosphere and radiative transfer code to synthesize the
spectrum. Unfortunately, it is computationally expensive to accurately generate
a spectrum over a wide range of wavelengths, and so these codes are generally
employed to generate libraries of stellar spectra, which exist over a range of
$\vg$ at specified points on a regular grid, typically covering the main sequence of stellar
types. Several high-quality libraries of synthetic spectra now exist
(\citealt{hwd+13}, Kurucz, and more are being developed for GAIA).

In a forward-modelling approach, which may make many calls to generate a
spectra of varying $\vg$, it is currently computationally impractical to
iterate the structure of the stellar atmosphere and then run the stellar
synthesis for each call of $\vg$.  Some methods compromise by taking a halfway
approach, whereby a precomputed library of stellar atmospheres is interpolated
for a given $\vg$, and then a simple radiative transfer code is run to
synthesize a spectrum (e.g.  SME; \citealt{vp96}). This technique has the
advantage that interpolating atmospheres and then running the spectral
synthesis for each $\vg$ might be more accurate than directly interpolating
synthetic spectra. However, if a large region of wavelength is required at high
resolution, then the synthesis step will be slow.  A different approach might
be instead to eschew forward modelling (and thus repeated synthesis or grid
interpolation) of the spectra entirely and evaluate the posterior probability
density only at the grid points of the spectral library.  Then, these discrete
samples of the posterior density can be interpolated to find the posterior
surface and draw confidence intervals on the parameters (e.g. SPC;
\citealt{blj+12}). However, the difficulty with this approach is that the
parameter uncertainty might be smaller than the grid spacing, and then
interpolation will not properly capture degeneracies between parameters.

Because we desire a generative model of the spectra without the added
complexity or computational overhead of running a stellar synthesis, we opt to
use a \todo{Implement a more accurate but still fast interpolator for the grid.
Likely this will be a combination of using splines to pre-interpolate a finer
grid (~20 K, 0.1 dex spacing in gravity and metallicity and then doing linear
interpolation off of this new grid during the actual MCMC run. It will be described here.} However, the
techniques we develop in this paper to account for and parameterize the
correlated, systematic pixel residuals are generally applicable to spectra fit
by any of these methods.

\begin{figure*}[!htb]
\begin{center}
  \includegraphics[width=0.85\textwidth]{figs/model_data.pdf}
  \caption{\textbf{Top}: Chebyshev polynomial which multiplies the model spectrum
   to account for inaccuracies in the flux-calibration. 
  \textbf{Middle}: The data spectrum and model spectrum, after it has been
   interpolated, post-processed, and multiplied by the Chebyshev polynomial.
  \textbf{Bottom}: Residuals from the model fit. Note the large residual at
   $\lambda5189$\AA\ due to a missing opacity source from \ion{Ca}{1}.}
\label{fig:model_data}
\end{center}
\end{figure*}

\subsection{Post-Processing}
\label{subsec:postprocess}

Libraries of synthetic spectra are generally computed at high resolution ($R
 \geq $100,000), sampled with many pixels per resolution element, and do not
 include any rotational broadening or account for any instrumental effects. 
In order to transform a raw synthetic spectrum interpolated from the grid into
 one that matches the spectrum of a real star, we must post-process the spectrum
 to account for these secondary effects. 
In addition to these fundamental parameters, a star also has several observed
 properties that are a function of its kinematics, geometry, and location in our
 galaxy: projected rotational velocity $v \sin i$, line of sight velocity $v_z$,
 total extinction $A_V$, and solid angle $\Omega$. 
We call these secondary parameters $\vpp$, because we can model their effects
 on the synthetic spectra in a ``post-processing'' step by convolution with an
 appropriate kernel or multiplication with a smooth function. 
Together, we call $\vstar = \{\vg, \vpp\}$.

The projected rotational velocity of the star, parameterized by $v \sin i$,
broadens spectral line profiles and is mathematically described by convolution
with a specific kernel \citep[Equation 18.14]{gra08}. UV, optical, and infrared
spectra are acquired using a spectrograph which disperses light onto a CCD,
defining a specific resolution and sampling rate for the spectrum.  Because the
resolution of the spectrograph and pixel sampling are different from the
synthetic spectral library, we must convolve the raw spectrum with the line
spread function (LSF) of the instrument and resample to the exact pixels of the
CCD. In wavelength space, the $v \sin i$ and LSF operations are represented by
the convolution of the synthetic spectrum with these two kernels.  Using the
convolution theorem, we can rewrite these convolutions as multiplications in
Fourier space.  In order to use the fast Fourier transform (FFT) with discrete
samples of the synthetic spectrum, we must first resample it
onto a uniform grid.  In the case of a spectrum, the natural
uniform grid is one that is equally sampled in velocity space, such there is an
equal velocity shift $\Delta v$ between pixels. This results in a wavelength
grid that is linearly sampled in the logarithm of wavelength.  Next, we
multiply the Fourier transform of the synthetic spectrum by the Fourier
transforms of the rotational velocity and line spread function kernels.
Lastly, we do an inverse FFT to transform the modified spectrum back to
wavelength space, where it is sampled at the wavelength locations of the pixels
in the detector.  When resampling the spectrum, we are careful to ensure
band-limited interpolation using splines in order to prevent introducing
non-physical high frequency structure into the spectrum.

We scale the raw synthetic flux (often reported at the stellar surface) to an
observed flux by multiplying by the solid angle of the star, $\Omega =
R_\star^2/d^2$, where $R_\star$ is the stellar radius and $d$ is the distance
to the star from earth. Then, we correct for a radial velocity difference $v_z$ between the
star and the earth by Doppler shifting the model spectrum.  Finally, we correct
for interstellar extinction by dividing by a dust extinction model
$A_\lambda$, which is parameterized by extinction in the $V$ photometric band,
$A_V$. 
\todo{We can add in convolution with photometric filter
profiles to use photometry and provide a better constraint on the luminosity of
the star}
\todo{We might try implementing the convolution in wavelength space, rather
 than the FFT. 
This would allow the LSF to change with wavelength.\\}

\paragraph{Calibration polynomials} In contrast to a collection of photometric points in a spectral energy distribution, or any ensemble of data for which each data point is acquired by an individual measurement, adjacent wavelength points in a spectrum are measured simultaneously, with the same detector and--more importantly--the same calibration. Typically, spectra are flux calibrated by dividing by a ``sensitivity function,'' which is determined by comparing an observation of a spectrophotometric standard to a previously calibrated model spectrum. Therefore, any error in the sensitivity function, which is typically a smooth function with wavelength, will create a region of the data spectrum which is systematically offset by a small amount from what the true flux should be. The flux calibration error is then the residual difference in the empirically determined sensitivity function from the what the ``true'' sensitivity function would be if we had a perfectly calibrated instrument and spectrophotometric standard. Therefore, it is necessary to introduce additional degrees of freedom into our model in order to account for the calibration error and allow the model spectrum to properly match the data. By adding the residual calibration as a parameter in our model, we can explore the distribution of possible calibrations (\S\ref{subsec:MCMC}) and then marginalize the calibration parameters out of the final posterior distribution of the stellar parameters. This means that we can derive estimates of stellar parameters that properly account for the intrinsic uncertainty in the flux calibration.

A traditional way to deal with errors in the flux calibration is to avoid flux-calibration entirely and simply normalize both the data and the spectrum to the continuum level before comparison.  While this procedure may work well for hotter stars with well-defined continuua, normalizing spectra of cooler stars with large molecular features may introduce artificial features into the spectrum when the pseudo-continuum is incorrectly placed.  Instead, we choose to parameterize the residual flux calibration error by a low-order Chebyshev polynomial which is normalized to unity. The coefficients of the polynomial are parameters in the model that will be determined by the fitting algorithm. If we have multiple observations of spectrophotometric standards, then we can characterize the typical flux-calibration error of the spectrograph and put priors on the coefficients that will prevent overfitting of the polynomials.  For an example of the Chebyshev polynomial determined for a typical spectrum, see Figure~\ref{fig:model_data}. Using priors on the Chebyshev coefficients ensures that real spectral features, such as deep molecular bands in later type stars, are not accidentally treated as calibration artifacts. The overall normalization of the polynomial $c_0$ is degenerate with the solid angle of the star $\Omega$. Therefore we apply an additional constraint and enforce that the mean of the Chebyshev polynomial be equal to $1$. If there is only one order in the spectrum, such as with a long-slit spectrum or a single order of an echelle spectrum, then this means setting $c_0 = 1$. If we are fitting more than one order of an echelle spectrum then we adopt an independent Chebyshev polynomial for each order. This means that we chose the $c_0$ of a specific order to anchor to 1, while the $c_0$ for other orders are allowed to take on perturbations about $1$.

This polynomial formalism can in principle be used to model raw (not flux calibrated) spectra by using the Chebyshev polynomial to solve directly for the sensitivity function itself, rather than a residual calibration error as is currently done. When used in this manner, there will be greater uncertainty about the properties of the spectrograph, which will translate into wider priors on the Chebyshev coefficients.  In this case, there is more of a chance that the polynomial might destroy broad-scale information present in the spectrum by fitting it out, something that would be avoided with more accurate priors provided by a flux calibration.

\subsection{Likelihood Calculation}
\label{subsec:likelihood}

To evaluate which parameters of the model spectrum fit the data spectrum best,
we use a likelihood function to compare individual pixels between the two
spectra. A likelihood function provides the probability of the data spectrum
$\fD$, given the model spectrum $\fM(\vt)$, which is parameterized by the
stellar and calibration parameters $\vt$. The data spectrum and model spectrum are both composed of $N$ pixels, and so $\fD$ and $\fM$ represent the collection of these pixels in vector form, 
$\fD = \{\fDi\}^N_{i=1}$ and $\fM = \{\fMi\}^N_{i=1} $, respectively. The vector of residuals is the difference between the data spectrum and the model spectrum
\begin{equation}
  \fR(\vt) = \fD - \fM(\vt)
\end{equation}
In the language of probability calculus, the likelihood function is written as
$p(\fD | \fM(\vt))$, or $p(\fD | \vt)$ for short. We choose a standard
multidimensional Gaussian likelihood function which allows for covariance
between the pixel residuals 
\begin{equation}
  p(\fD | \vt) = \frac{1}{\sqrt{(2 \pi)^N \det(C)}} \exp\left ( -\frac{1}{2}
   \fR^T C^{-1} \fR \right ) 
   \label{eqn:prob}
\end{equation}
which penalizes models that yield larger residuals. $C$ is the covariance
matrix which describes the covariance between the pixel residuals. The structure of the covariance matrix is discussed in detail in the next section (\S\ref{subsec:covariance}). Because
the likelihood function generally exhibits a large dynamic range, it is numerically
preferred to use the log-likelihood function
\begin{equation}
  \ln[p(\fD | \vt)] = -\frac{1}{2} \fR^T C^{-1} \fR - \frac{1}{2} \ln \det C  
   - \frac{N}{2} \ln 2 \pi 
  \label{eqn:lnprob}
\end{equation}

\subsection{Parameterizing the covariance structure}
\label{subsec:covariance}
The covariance matrix $C$ describes the measurement noise in each pixel of the spectrum as well as the covariance between pixel residuals. It is fundamental to the likelihood calculation described in \S\ref{subsec:likelihood}. For most problems, such as fitting a spectral energy distribution to photometric measurements, each data point is acquired in an independent measurement from the other data points, which means that the noise for each data point, $\sigma_i$, is also independent from the other data points. This results in a diagonal covariance matrix
\begin{equation}
  C = 
  \begin{bmatrix}
    \sigma_0^2 & 0  & \cdots & 0\\
    0 & \sigma_1^2 & \cdots & 0\\
    \vdots  &   & \ddots  & \vdots \\
    0 & 0 & \cdots & \sigma_N^2\\
  \end{bmatrix}
  \label{eqn:covariance_diagonal}
\end{equation}
with $\sigma_{ij} = \delta_{ij} \sigma_{ij}$, where $\delta_{ij}$ is the Kronecker delta function. When the covariance matrix is diagonal, 
the likelihood function (Equation~\ref{eqn:lnprob}) reduces to the familiar
$\chi^2$ form of a sum over the square of the residuals, weighted by the
inverse variance of each data point
\begin{equation}
  \ln[p(\fD | \vt)] \propto - \frac{1}{2} \chi^2 = - \frac{1}{2} \sum_i^N
   \frac{R_i^2}{\sigma_i^2}
   \label{eqn:chi2}
\end{equation}

\begin{figure*}[!htb]
\begin{center}
  \includegraphics{figs/badlines.pdf}
  \caption{A collection of spectral lines induced by imperfect model fits.
    From left to right: \textbf{Class 0} The majority of spectral lines
    ($\gtrsim 60$\%) will have minor differences in strength between data and
    model spectrum, which produce low-amplitude correlations in the residuals
    on the length scale of the width of a typical spectral line.  \textbf{Class
    I}: Sometimes ($\lesssim 5$\% of all lines), a missing opacity source in
    the model (in this case a line-blended \ion{Ca}{1}) leaves a large, highly correlated
    patch of negative residuals.  \textbf{Class II}: Sometimes ($\lesssim 5$\%
    of all lines), an extraneous line in the model leaves a large, highly
    correlated patch of positive residual.  \textbf{Class III}: If the line strengths are
    substantially discrepant ($\lesssim 10$\% of all lines), there will be many
    correlated residuals of moderate amplitude.  The difficulty
    with class III lines is that for any specific line, there might exist a
    $\vstar$ that will fit the line, but there does not exist a $\vstar$ that
    will properly fit \emph{all} the lines.}
\label{fig:badlines}
\end{center}
\end{figure*}

However, the pixel residuals from a spectroscopic fit will often be correlated and will invalidate Equation~\ref{eqn:chi2}. These correlations arise because the CCD pixels of a spectrograph do not sample independent wavelengths of the spectrum. Instead, spectrographs are designed so that the LSF is properly sampled by at least a few pixels, meaning that adjacent pixels do not sample completely independent parts of the spectrum. Because of this, when a spectral feature differs between the model and the data it creates a residual which spans several adjacent data points. For a collection of correlated residuals produced by a typical spectroscopic fit, see Figure~\ref{fig:badlines}.
